---
title: "ANOVA Project"
author: "abdellah"
date: "2023-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(nlme)
library(multcomp)
library(tidyverse)
library(ggplot2)
library(car)
```



```{r}
# Load the data
Wastedf<- Wastedata1 %>%
  filter(County==c("Hennepin","Ramsey","Anoka")) %>%
  filter(Category==c("Paper","Metal","Organic")) #%>%
  #select(County, Category, `Res Tons`)
Wastedf$County <-as.factor(Wastedf$County)
Wastedf$Category<-as.factor(Wastedf$Category)
Wastedf
# Run the ANOVA analysis
model1 <- lm(`Res Tons`~ County + Category + County*Category, data = Wastedf)
Anova(model1, type = "II")
# interaction analysis
paper <- Wastedf %>%
  filter((Category == "Paper"))
paperModel<- aov(`Res Tons`~County, data = paper)
organic <- Wastedf %>%
  filter((Category == "Organic"))
organicModel<- aov(`Res Tons`~County, data = organic)
metal <- Wastedf %>%
  filter((Category == "Metal"))
metalModel<- aov(`Res Tons`~County, data = metal)
summary(paperModel)
summary(organicModel)
summary(metalModel)
# Calculate avg tons per county and category, then graph
Wastedf %>%
  group_by(County)%>%
  summarize(aveWaste=mean(`Res Tons`)) %>%
  ggplot(aes(x = County, y = aveWaste)) + geom_bar(stat = "identity", fill = "darkblue")+labs(title="Mean Tons of Recycled Waste Among Counties", x = "Counties", y = "Mean Recycled Waste (Tons)")

Wastedf %>%
  group_by(Category) %>%
  summarize(aveWaste = mean(`Res Tons`)) %>%
  ggplot(aes(x = Category, y = aveWaste)) + geom_bar(stat = "identity", fill = "darkgreen")+labs(title="Mean Tons of Recycled Waste Among Categories", x = "Category", y = "Mean Recycled Waste (Tons)")


```

```{r}
#sessionInfo()
dfWaste2 <- Wastedata2
# Convert the year variable to a factor
dfWaste2$Year <- as.factor(dfWaste2$Year)
dfWaste2$County<- as.factor(dfWaste2$County)
#dfWaste2
dfWaste2 <- dfWaste2%>% 
  group_by(Year)%>% 
  select(Year, Landfilled, WTE)%>% 
  filter(Year==c(1991, 2017))%>% 
  
  gather(key=wasteType, value=wasteperTon, "Landfilled", "WTE")
dfWaste2
dfWaste2$wasteType<- as.factor(dfWaste2$wasteType)

# Run the two-way ANOVA model
model2 <- aov(wasteperTon~ Year+ wasteType + Year * wasteType, data = dfWaste2)

# Print the ANOVA table
summary(model2)

# interaction plot
interactionWaste2<- dfWaste2 
interaction.plot(interactionWaste2$wasteType,interactionWaste2$Year,interactionWaste2$wasteperTon, fun=mean, ylab = "Waste per Ton",xlab = "Waste Type",col = c("red", "blue","green"), trace.label = "Year")
```
A tibble:172 × 3
Groups:Year [2]
Year
<fctr>
wasteType
<chr>
wasteperTon
<dbl>
1991	Landfilled	0		
1991	Landfilled	15130		
1991	Landfilled	0		
1991	Landfilled	35531		
1991	Landfilled	6288		
1991	Landfilled	6584		
1991	Landfilled	11429		
1991	Landfilled	12		
1991	Landfilled	5437		
1991	Landfilled	158970		
...
1-10 of 172 rows
grouped_df
172 x 3
                Df    Sum Sq   Mean Sq F value Pr(>F)
Year             1 3.593e+09 3.593e+09   1.505  0.222
wasteType        1 2.874e+08 2.874e+08   0.120  0.729
Year:wasteType   1 3.281e+07 3.281e+07   0.014  0.907
Residuals      168 4.009e+11 2.386e+09               
R Console
A tibble:172 × 3
Groups:Year [2]
Year
<fctr>
wasteType
<fctr>
wasteperTon
<dbl>
1991	Landfilled	0		
1991	Landfilled	15130		
1991	Landfilled	0		
1991	Landfilled	35531		
1991	Landfilled	6288		
1991	Landfilled	6584		
1991	Landfilled	11429		
1991	Landfilled	12		
1991	Landfilled	5437		
1991	Landfilled	158970		
...
1-10 of 172 rows
grouped_df
172 x 3

A tibble:172 × 3
Groups:Year [2]
Year
<fctr>
wasteType
<fctr>
wasteperTon
<dbl>
1991	Landfilled	0		
1991	Landfilled	15130		
1991	Landfilled	0		
1991	Landfilled	35531		
1991	Landfilled	6288		
1991	Landfilled	6584		
1991	Landfilled	11429		
1991	Landfilled	12		
1991	Landfilled	5437		
1991	Landfilled	158970		
...
1-10 of 172 rows
grouped_df
172 x 3





